---
title: "Comparing Methods"
---

# Prepare environment

Packages

```{r}
library(tidyverse)
library(plotly)
library(googledrive)
library(readxl)
source("Code/Import_datasets.R")
'%!in%' <- function(x,y)!('%in%'(x,y))
```

Download nessecery data and update from google drive:
```{r}
# Confirm authorisation
drive_auth()

# Update datasets
update_datasets()

```

# 1. Import and combine Sequence datasets

## COI data components
```{r}

# Import downloaded rds files
ASV_COI <- readRDS("Data_import/COI/ASV_COI.rds")
Samptab_COI <- readRDS("Data_import/COI/Samptab_COI.rds")
Taxtab_COI <- readRDS("Data_import/COI/Taxtab_COI.rds")

# Combine data components to data frame
df_COI <- 
  Taxtab_COI %>%
  full_join(ASV_COI, by = "ASV") %>%
  pivot_longer(!1:12, names_to = "Library_ID", values_to = "Abundance") %>%
  full_join(Samptab_COI, by = "Library_ID")
```

## 18S data components

```{r}

# Import downloaded rds files
ASV_18S <- readRDS("Data_import/18S/ASV_18S.rds")
Samptab_18S <- readRDS("Data_import/18S/Samptab_18S.rds")
Taxtab_18S <- readRDS("Data_import/18S/Taxtab_18S.rds")

# Combine data components to data frame
df_18S <- 
  Taxtab_18S %>%
  full_join(ASV_18S, by = "Sequence") %>% 
  pivot_longer(!1:10, names_to = "Library_ID", values_to = "Abundance") %>% 
  full_join(Samptab_18S, by = "Library_ID") %>% 
  filter(is.na(ASV)==FALSE)

df_18S
```

## Microscopy data

```{r}
df_count_vertical <- read_excel("Data_import/Microscopy/Vertical_ZP_2022.xlsx")

df_count_2017 <- readRDS("Data_import/Microscopy/Hakai_Zooplankton_2013-2017_AN.rds") %>% 
  mutate(Name = Taxon) %>% 
  separate(Taxon, sep= " ", into=c("Genus", "Tax1", "Tax2", "Tax3", "Tax4"))
```


Specifying global filterine parameter:
What phyla are present in Metazoa?
```{r}
metazoa <- c("Arthropoda",
             "Cnidaria",
             "Chaetognatha",
             "Annelida",
             "Porifera",
             "Echinodermata",
             "Mollusca",
             "Phoronida",
             "Bryozoa",
             "Nemertea",
             "Ctenophora",
             "Priapulida",
             "Rotifera",
             "Chordata",
             "Platyhelminthes")
```


# 2. Vertical Comparaison eDNA index

## 2.1 Zooplankton net samples:

### Microscopy

```{r}
  
net_microscopy_index <-
  df_count_vertical %>% 
  mutate(Abundance=Count*Fraction) %>%
  
  # Summarise all Genus counts per
  group_by(StationID, Depth, SampleID, Genus) %>% 
  summarise(Abundance = sum(Abundance)) %>%
  
  # Fill zero counts
  pivot_wider(names_from = Genus,
              values_from = Abundance,
              values_fill = 0) %>% 
  pivot_longer(4:11, names_to = "Genus", values_to = "Abundance") %>% 
  
  # Calculate relative abundance
  group_by(StationID, Depth) %>% 
  reframe(Abundance = Abundance/sum(Abundance),
            Genus) %>% 
  
  # Calculate Index
  group_by(Genus) %>% 
  reframe(eDNA_index = Abundance/max(Abundance),
          Site_ID = StationID, Line_out_depth = Depth) %>% 
  mutate(Marker = "Microscopy")

net_microscopy_index
  
```
```{r}
counted_taxa_vertical <-
  net_microscopy_index %>% 
  pull(Genus) %>% 
  unique()

counted_taxa_vertical
```



### 18S

```{r}

Net_18S_index <-
  df_18S %>% 
  filter(Project_name == "Vertical",
         year(Sample_date) == 2022,
         Division == "Metazoa",
         Sample_type == "Zp_Net",
         is.na(Genus) == FALSE,
         Genus %in% counted_taxa_vertical) %>% 
  
  #sum over genus and remove low counts
  group_by(Genus, Line_out_depth, Site_ID) %>% 
  summarise(Abundance = sum(Abundance)) %>%
  group_by(Genus) %>% 
  filter(sum(Abundance) > 100) %>% 
  
  #relative abundance
  group_by(Line_out_depth, Site_ID) %>% 
  reframe(relative_abundance = Abundance / sum(Abundance, na.rm = TRUE),
          Genus) %>%  
  #eDNA index
  group_by(Genus) %>% 
  reframe(eDNA_index = relative_abundance/max(relative_abundance),
          Line_out_depth, Site_ID) %>%
  filter(is.na(eDNA_index) == FALSE) %>% 
  
  mutate(Line_out_depth = factor(Line_out_depth, levels = c(
    "80-0", "120-80", "200-120"))) %>% 
  mutate(Marker = "18S")

Net_18S_index
```

### COI

```{r}
Net_COI_index <- 
  df_COI %>%
  filter(Project_name == "Vertical",
         year(Sample_date) == 2022,
         phylum %in% metazoa,
         Sample_type == "Zp_Net",
         is.na(genus) == FALSE,
         genus %in% counted_taxa_vertical) %>%
  
  #sum over genus and remove low counts
  group_by(genus, Line_out_depth, Site_ID) %>% 
  summarise(Abundance = sum(Abundance)) %>%
  group_by(genus) %>% 
  filter(sum(Abundance) > 100) %>% 
  
  #relative abundance
  group_by(Line_out_depth, Site_ID) %>% 
  reframe(relative_abundance = Abundance / sum(Abundance, na.rm = TRUE),
          genus) %>%  
  #eDNA index
  group_by(genus) %>% 
  reframe(eDNA_index = relative_abundance/max(relative_abundance),
          Line_out_depth, Site_ID) %>%
  filter(is.na(eDNA_index) == FALSE) %>%
  
  mutate(Line_out_depth = factor(Line_out_depth, levels = c(
    "80-0", "120-80", "200-120")),
    Genus = genus,
    Marker = "COI") %>% 
  select(!genus)

Net_COI_index
```




### Combine and plot:

```{r}
comp_plot <-
  bind_rows(Net_COI_index, Net_18S_index, net_microscopy_index) %>%
  mutate(RA_Index = eDNA_index,
         Depth = factor(Line_out_depth, levels = c(
           "200-120", "120-80", "80-0"))) %>%
  ggplot() +
  geom_point(aes(RA_Index, Depth, shape = Site_ID, color = Marker)) +
  stat_summary(aes(RA_Index, Depth, color = Marker, group = Marker),
               fun=mean, geom="line") +
  facet_wrap(facets = "Genus") +
  theme_bw() +
  labs(title = "Zooplankton Net Samples - Vertical")

comp_plot
ggsave("Figure_output/ZPnet_Vertical.pdf")

```

## 2.2 eDNA Samples

### 18S

```{r}
eDNA_18S_index <- 
df_18S %>% 
  filter(Project_name == "Vertical",
         year(Sample_date) == 2022,
         Division == "Metazoa",
         Sample_type == "eDNA",
         is.na(Genus) == FALSE) %>% 
  
  #sum over genus and remove low counts
  group_by(Genus, Line_out_depth, Site_ID) %>% 
  summarise(Abundance = sum(Abundance)) %>%
  group_by(Genus) %>% 
  filter(sum(Abundance) > 100) %>% 
  
  #relative abundance
  group_by(Line_out_depth, Site_ID) %>% 
  reframe(relative_abundance = Abundance / sum(Abundance, na.rm = TRUE),
          Genus) %>%  
  
  #eDNA index
  group_by(Genus) %>% 
  reframe(eDNA_index = relative_abundance/max(relative_abundance),
          Line_out_depth, Site_ID) %>%
  filter(is.na(eDNA_index) == FALSE) %>% 
  mutate(Line_out_depth = as.integer(Line_out_depth),
         Marker = "18S")

eDNA_18S_index
```

### COI

```{r}
eDNA_COI_index <- 
  df_COI %>%
  filter(Project_name == "Vertical",
         year(Sample_date) == 2022,
         phylum %in% metazoa,
         Sample_type == "eDNA",
         is.na(genus) == FALSE) %>%
  
  #sum over genus and remove low counts
  group_by(genus, Line_out_depth, Site_ID) %>% 
  summarise(Abundance = sum(Abundance)) %>%
  group_by(genus) %>% 
  filter(sum(Abundance) > 100) %>% 
  
  #relative abundance
  group_by(Line_out_depth, Site_ID) %>% 
  reframe(relative_abundance = Abundance / sum(Abundance, na.rm = TRUE),
          genus) %>%  
  #eDNA index
  group_by(genus) %>% 
  reframe(eDNA_index = relative_abundance/max(relative_abundance),
          Line_out_depth, Site_ID) %>%
  filter(is.na(eDNA_index) == FALSE) %>% 
  
  mutate(Line_out_depth = as.integer(Line_out_depth),
         Marker = "COI",
         Genus = genus) %>% 
  select(!genus)

eDNA_COI_index

```

### Combine and Plot

```{r}
eDNA_plot <- bind_rows(eDNA_COI_index, eDNA_18S_index) %>%
  ggplot() +
  geom_point(aes(Line_out_depth, eDNA_index, shape = Site_ID, color = Marker)) +
  stat_summary(aes(Line_out_depth, eDNA_index, color = Marker, group = Marker),
               fun=mean, geom="line") +
  facet_wrap(facets = "Genus") +
  theme_bw() +
  labs(title = "eDNA samples")

eDNA_plot
ggsave("Figure_Output/eDNA_vertical.pdf")
```



# 3. Temporal Comparaison

## 3.1 Zoop Net tows

### Microscopy

```{r}


Net_count_index_2017 <-
  df_count %>% 
  filter(Station %in% c("QU39", "QU24")) %>% 
  mutate(Biomass=ifelse(is.na(Biomass), 0, Biomass)) %>% 
  filter(year(Date) == 2017) %>% 
  group_by(Date, Genus) %>%
  #filter(Genus %in% COI_genus) %>% 
  summarize(Biomass = sum(Biomass),
            Depth=as.numeric(DEPTH_STRT)) %>% 
  mutate(Biomass = Biomass*Depth) %>% 
  filter(Biomass > 10) %>% 
  group_by(Date, Genus) %>% 
  summarise(Abundance = sum(Biomass)) %>% 
  
  # Calculate relative abundance
  group_by(Date) %>% 
  reframe(RelativeAbundance=Abundance/sum(Abundance),
          Genus) %>%
  filter(is.na(RelativeAbundance) == FALSE) %>%
  
  group_by(Genus) %>%
  reframe(RA_Index = RelativeAbundance/max(RelativeAbundance),
          Date) %>%
  
  filter(is.na(RA_Index) == FALSE) %>%
  group_by(Genus) %>%
  filter(n() > 3) %>% 
  mutate(Marker = "Microscopy")
  

Net_count_index_2017
Net_count_index_2017 %>%
  ggplot() +
  geom_line(aes(Date, RA_Index)) +
  facet_wrap(facets = "Genus")
  

counted_taxa <- filt_count %>% 
pull(Genus) %>% unique

counted_genera <- Net_count_index_2017 %>% 
  pull(Genus) %>%  unique
counted_genera
```
### COI

```{r}
Net_COI_index_2017 <- 
  df_COI %>%
  filter(Project_name == "Vertical",
         year(Sample_date) == 2017,
         phylum %in% metazoa,
         Sample_type == "Zp_Net",
         genus %in% counted_genera) %>% 
  
  #sum over genus and remove low counts
  group_by(genus, Sample_date, Site_ID) %>% 
  summarise(Abundance = sum(Abundance)) %>%
  group_by(genus) %>% 
  filter(sum(Abundance) > 100) %>% 
  
  #relative abundance
  group_by(Sample_date, Site_ID) %>% 
  reframe(relative_abundance = Abundance / sum(Abundance, na.rm = TRUE),
          genus) %>%  
  #eDNA index
  group_by(Genus = genus) %>% 
  reframe(RA_Index = relative_abundance/max(relative_abundance),
          Date = date(Sample_date),
          Marker = "COI") %>%
  filter(is.na(RA_Index) == FALSE)

Net_COI_index_2017

Net_COI_index_2017 %>% 
  ggplot() +
  geom_line(aes(Date, RA_Index)) +
  facet_wrap(facets = "Genus")
```

### 18S

```{r}
Net_18S_index_2017 <-
  df_18S %>% 
  filter(Project_name == "Vertical",
         year(Sample_date) == 2017,
         Division == "Metazoa",
         Sample_type == "Zp_Net",
         is.na(Genus) == FALSE,
         Genus %in% counted_genera) %>% 
  
  #sum over genus and remove low counts
  group_by(Genus, Sample_date, Site_ID) %>% 
  summarise(Abundance = sum(Abundance)) %>%
  group_by(Genus) %>% 
  filter(sum(Abundance) > 100) %>% 
  
  #relative abundance
  group_by(Sample_date, Site_ID) %>% 
  reframe(relative_abundance = Abundance / sum(Abundance, na.rm = TRUE),
          Genus) %>%  
  #eDNA index
  group_by(Genus) %>% 
  reframe(RA_Index = relative_abundance/max(relative_abundance),
          Date = date(Sample_date),
          Marker = "18S") %>%
  filter(is.na(RA_Index) == FALSE) %>% 
  
  mutate(Marker = "18S")

Net_18S_index_2017

Net_18S_index_2017 %>% 
  ggplot() +
  geom_line(aes(Date, RA_Index)) +
  facet_wrap(facets = "Genus")
 
```




### Combine plots

```{r}
bind

Net_count_index_2017
Net_COI_index_2017
Net_18S_index_2017


comp_plot_2017 <-
  bind_rows(Net_count_index_2017,
            Net_COI_index_2017,
            Net_18S_index_2017) %>% 
  ggplot() +
  geom_line(aes(Date, RA_Index, color = Marker)) +
  facet_wrap(facets = "Genus") +
  theme_bw() +
  labs(title = "Integrated Zooplankton Nets Seasonal") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

comp_plot_2017
ggsave("Figure_Output/ZPnet_Seasonal.pdf")

```




## 3.2 eDNA

### COI

```{r}
filt_COI <- df_COI %>% 
  filter(Project_name == "QU39",
         Sample_type == "eDNA",
         year(Sample_date)==2017,
         phylum %in% metazoa,
         genus %in% counted_genera) %>% 
  
  #sum over genus and remove low counts
  group_by(genus, Line_out_depth, Sample_date) %>% 
  summarise(Abundance = sum(Abundance)) %>%
  ungroup() %>%  
  
  group_by(genus) %>%

  
  #relative abundance
  group_by(Line_out_depth, Sample_date) %>% 
  reframe(relative_abundance = Abundance / sum(Abundance),
          Genus = genus, Abundance) %>%
  filter(is.na(relative_abundance) == FALSE) %>%
  #eDNA index
  group_by(Genus) %>% 
  reframe(eDNA_index = relative_abundance/max(relative_abundance),
          Abundance = Abundance/max(Abundance),
          Line_out_depth, Sample_date, ) %>% 
  filter(is.na(eDNA_index) == FALSE) %>% 
    mutate(Line_out_depth = factor(Line_out_depth, levels = rev(c(
    "0", "5", "30", "100", "260"))),
    Marker = "COI")
filt_COI

eDNA_plot_2017 <- 
  filt_COI %>% 
  mutate(Depth = factor(Line_out_depth, levels = rev(c(
    "0", "5", "30", "100", "260")))) %>%
  #mutate(Depth = as.numeric(Line_out_depth)) %>% 
  ggplot() +
  geom_point(aes(Sample_date, Depth, color = eDNA_index, size = eDNA_index, alpha = Abundance)) +
  facet_wrap(facets = "Genus") +
  theme_bw() +
  scale_color_viridis_b() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "eDNA Seasonal and Vertiva")

eDNA_plot_2017
ggsave("Figure_Output/eDNA_2017.pdf")

  


```

### 18S

```{r}
ps_18S <- readRDS("ProcessedData/18S_Caterina/ps_18S.rds") %>% psmelt()

filt_18S <-
  ps_18S %>% 
  filter(year(collected) == 2017,
         Genus %in% counted_genera,
         site_id == "QU39") %>% 
  mutate(Sample_date = date(collected)) %>% 
  
  #sum over genus and remove low counts
  group_by(Genus, line_out_depth, Sample_date) %>% 
  summarise(Abundance = sum(Abundance)) %>%
  ungroup() %>% 
  
  #group_by(Genus) %>% 
  #filter(any(max(Abundance) > 100),
  #       any(sum(Abundance) > 500)) %>% 

  
  #relative abundance
  group_by(line_out_depth, Sample_date) %>% 
  reframe(relative_abundance = Abundance / sum(Abundance),
          Genus) %>%
  filter(is.na(relative_abundance) == FALSE) %>% 
  #eDNA index
  group_by(Genus) %>% 
  reframe(eDNA_index = relative_abundance/max(relative_abundance),
          Line_out_depth = line_out_depth, Sample_date) %>% 
  filter(is.na(eDNA_index) == FALSE) %>% 
  mutate(Line_out_depth = factor(Line_out_depth, levels = rev(c(
    "0", "5", "30", "100", "260"))),
    Marker = "18S")
filt_18S


eDNA_plot_2017 <- 
  filt_18S %>% 
  mutate(Depth = factor(Line_out_depth, levels = rev(c(
    "0", "5", "30", "100", "260")))) %>% 
  #mutate(Depth = as.numeric(Line_out_depth)) %>% 
  ggplot() +
  geom_point(aes(Sample_date, Depth, color = eDNA_index, size = eDNA_index, alpha = eDNA_index)) +
  facet_wrap(facets = "Genus") +
  theme_bw() +
  scale_color_viridis_b() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "eDNA Seasonal and Vertiva")
  
eDNA_plot_2017

```

### Merge plots

```{r}
bind_rows(filt_COI, filt_18S) %>% 
  group_by(Genus, Line_out_depth, Sample_date) %>% 
  summarise(eDNA_index = mean(eDNA_index)) %>%
  ggplot() +
  geom_point(aes(Sample_date, Line_out_depth, color = eDNA_index, size = eDNA_index, alpha = eDNA_index)) +
  facet_wrap(facets = "Genus") +
  theme_bw() +
  scale_color_viridis_b() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "eDNA Seasonal and Vertiva")
  
  
  bind_rows(filt_COI, filt_18S) %>% 
  group_by(Genus, Line_out_depth, Sample_date) %>% 
  summarise(eDNA_index = mean(eDNA_index)) %>% 
  pull(Genus) %>% unique
```



